PAMELA: an annotation-based Java Modeling Framework

### 2021-02-17 related.tex
related after Antidote

The PAMELA framework can be seen as a CASE tool that focuses in the design and verification and validation phases of the software development life-cycle. In that sense, it presents similarities to other existing object oriented CASE tools and/or language workbenches such as Kermeta~\cite{kermeta} and the Platypus~\cite{platypus} meta-case tool as they permit the definition and manipulation of models, metamodels and constraints/behavior. Two main features separate PAMELA from those approaches though: 1) The main artefact for the PAMELA framework is standard java code; and 2) PAMELA does not rely on code generation but on the interpretation/construction of models at runtime guided by code annotations. This allows for the seamlessly blending of the metamodeling/programming phases.

When the observed artifact is code (e.g., for runtime analysis or verification of properties) we could see our approach as similar to some contributions appertaining to the models@runtime topic~\cite{bencomo2019models}. Indeed models@runtime approaches often rely  on the use of the reflection architectural pattern~\cite{buschmann2008pattern} in order to separate core application logic from a metalevel that contains information about properties, types, etc. This is also the case of PAMELA. 
%However the focus of most models@runtime approaches is not on easing the development process but on enabling the analysis and/or the adaptation/evolution of ever running applications. 
Examples of such approaches are: AC-Contrat~\cite{accontract} which provides runtime verification of properties for context-aware applications; Ramses~\cite{ramses} focused on dynamic adaptations; and \cite{denker2010modeling} that performs feature analysis.~\cite{song2010applying} wraps running systems in standard UML-like models in order to perform analysis and management tasks by using off-the-shelf MDE tools and techniques whereas FAME \cite{kuhn2008fame} is a polyglot library that keeps metamodels accessible and adaptable at runtime (Synchronization of changes between models and code are however limited for languages such as Java).

@@

More similar to PAMELA (as it focus on avoiding round-tripping issues) in~\cite{cavarle2016dynamic} the authors present an approach to keep the (bi-directional) synchronization between feature-based models and generated code alternatives. They rely on Pharo\cite{pharo} reflective capabilities. 
More recently, in \cite{boronat2019} the authors construct and maintain at runtime model-based views on the data manipulated in object-oriented code.

>>>>

The PAMELA framework can be seen as a CASE tool that focuses on the design and verification and validation phases of the software development life-cycle. In that sense, it presents similarities to other existing object oriented CASE tools and/or language workbenches such as Kermeta~\cite{kermeta} and the Platypus~\cite{platypus} meta-case tool as they permit the definition and manipulation of models, metamodels and constraints/behavior. Two main features separate PAMELA from those approaches though: 1) The main artifact for the PAMELA framework is standard java code; and 2) PAMELA does not rely on code generation but on the interpretation/construction of models at runtime guided by code annotations. This allows for the seamlessly blending of the metamodeling/programming phases.

When the observed artifact is code (e.g., for runtime analysis or verification of properties) we could see our approach as similar to some contributions appertaining to the models@runtime topic~\cite{bencomo2019models}. Indeed models@runtime approaches often rely  on the use of the reflection architectural pattern~\cite{buschmann2008pattern} in order to separate core application logic from a metalevel that contains information about properties, types, etc. This is also the case of PAMELA. 
%However the focus of most models@runtime approaches is not on easing the development process but on enabling the analysis and/or the adaptation/evolution of ever running applications. 
Examples of such approaches are: AC-Contrat~\cite{accontract} which provides runtime verification of properties for context-aware applications; Ramses~\cite{ramses} focused on dynamic adaptations; and \cite{denker2010modeling} that performs feature analysis.~\cite{song2010applying} wraps running systems in standard UML-like models in order to perform analysis and management tasks by using off-the-shelf MDE tools and techniques whereas FAME \cite{kuhn2008fame} is a polyglot library that keeps metamodels accessible and adaptable at runtime (Synchronization of changes between models and code are, however, limited for languages such as Java).

@@

More similar to PAMELA (as it focus on avoiding round-tripping issues) in~\cite{cavarle2016dynamic} the authors present an approach to keep the (bidirectional) synchronization between feature-based models and generated code alternatives. They rely on Pharo\cite{pharo} reflective capabilities. 
More recently, in \cite{boronat2019} the authors construct and maintain at runtime model-based views on the data manipulated in object-oriented code.
